Formatting Instructions for ICLR 2020 \\ Conference Submissions

### 2019-09-10 reports/iclr/paper.tex
Update to ICLR notation
Add table compare attentions + restructure

\section{Attention Layer Can Implement CNN}

@@

We take inspiration from \citep[TransformerXL]{dai2019transformerxl} to define relative positional encoding.
They decompose the computation of the attention coefficients as follow:
\begin{align}
    \mathbf{A}_{i, j}^{\mathrm{abs}} &= (\mE_{x_i} + \mU_i)^\top \mW_q^\top \mW_k (\mE_{x_j} + \mU_j)\\
  %
  %
 &=
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{E}_{x_{j}}}_{(a)} 
  +
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{U}_{j}}_{(b)}
  +
  \underbrace{\mathbf{U}_{i}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{E}_{x_{j}}}_{(c)}
  +
  \underbrace{\mathbf{U}_{i}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{U}_{j}}_{(d)}
\end{align}
with $\mE_{x_i}$ the embedding of token $x_i$, $\mU_i$ the absolute position encoding, $\mR_{i-j}$ relative position encoding. They replace all absolute position encoding with relative ones:
\begin{align}
  \mathbf{A}_{i, j}^{\mathrm{rel}} &=
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k, E} \mathbf{E}_{x_{j}}}_{(a)}
  +
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k, R} {\color{blue} \mathbf{R}_{i-j}}}_{(b)}
  +
  % \underbrace{{\color{red} \vu^{\top}} \mathbf{W}_{k, E} \mathbf{E}_{x_{j}}}_{(c)}
  +
  \underbrace{{\color{red} \vv^{\top}} \mathbf{W}_{k, R} {\color{blue} \mathbf{R}_{i-j}}}_{(d)}
  \label{eq:att_rel}
\end{align}

@@

\begin{align}
  f_{\mu, \Sigma}(\vx) = \exp \left(
    - \frac 1 2
    (\vx - \mu)^\top \Sigma^{-1} (\vx - \mu)
  \right)
\end{align}
We drop the normalization factor of the Gaussian distribution as it is replaced by the softmax of the attention mechanism.
The matrix $\Sigma^{-1}$ must be positive semi-definite.
We parametrize it as 
\begin{align}
  \Sigma^{-1} = (\Sigma^{-1/2})^\top\Sigma^{-1/2} = \begin{pmatrix}a & b \\ b & c\end{pmatrix}  
\end{align}
Our goal is to write the exponent as a dot product of two vectors:
\begin{itemize}
  \item one vector $R_\vx$ dependent only on $\vx$, the relative positional encoding,
  \item one vector $\mathbf{t}_h$ dependent only on $\mu$ and $\Sigma$, the attention head.
\end{itemize}
\begin{align}
  R_\vx &= 
  \begin{pmatrix}
    x_1      & x_2       & x_1      & x_2      & x_1^2   & x_2^2 & x_1 x_2 & 1        & 1         & 1
  \end{pmatrix}\\
  \mathbf{t}_h &= -\frac{1}{2}
  \begin{pmatrix}
    -2a\mu_1 & -2c \mu_2 & -2b\mu_2 & -2b\mu_1 & a       & c     & 2b      & a\mu_1^2 & c\mu_2^2 & 2b\mu_1 \mu_2
  \end{pmatrix}\\
  R_\vx^\top \mathbf{t}_h &=
  -\frac{1}{2} 
  \left[
  a(x_1 - \mu_1)^2 + 2b(x_1-\mu_1)(x_2 -\mu_2) + c(x_2-\mu_2)^2 
  \right]
  = - \frac 1 2 (\vx - \mu)^\top \Sigma^{-1} (\vx - \mu)
\end{align}
We can check that setting $b = 0$ and $a=c=\alpha$ recovers the expression of the isotropic Gaussian derived above.

\subsection{Extensions of the Attention Mechanism?} 
\begin{itemize}

@@

\begin{figure}[h!]
\includegraphics[width=.16\linewidth]{plots/attention/layer_0.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_1.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_2.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_3.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_4.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_5.png}
\caption{Contours of attention weights per layer for the 6 layers fully attentional model.}
\end{figure}

\begin{figure}[h!]
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_0.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_1.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_2.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_3.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_4.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_5.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_6.png}
\caption{Contours of attention weights during training (300 epochs) at layer 3.}
\end{figure}

@@

\section{Convolutional Layer}

\subsubsection*{Author Contributions}
If you'd like to, you may include  a section for author contributions as is done
in many journals. This is optional and at the discretion of the authors.

@@

\appendix
\section{Appendix}
You may include other additional sections here.

>>>>

\section{Introduction}

\section{Background on Attention Mechanism and Related Work}

\subsection{Convolution Layer}

\subsection{Self Attention Layer}

\subsection{Attention Mechanism in Vision}

Successful work to use attention on images \citep{ramachandran2019standaloneselfattention,belloAttentionAugmentedConvolutional2019}.

\begin{table}[h]
  \centering
  \begin{tabular}{lccc}
    \toprule
    Model&relative&sinusoids&learned\\
    \midrule
    \cite{vaswani17attentionisallyouneed} & & \checkmark\\
    \cite{radford2018gpt2} & & & \checkmark\\
    \cite{devlin2018bert} & & & \checkmark\\
    \cite{dai2019transformerxl} & \checkmark & \checkmark \\
    \cite{yang2019xlnet}  & \checkmark & \checkmark \\
    \midrule
    \cite{belloAttentionAugmentedConvolutional2019} & \checkmark & & \checkmark \\
    \cite{ramachandran2019standaloneselfattention} & \checkmark & & \checkmark \\
    % Images
    \bottomrule
  \end{tabular}
  \caption{Type of position encoding used by transformers models applied to text (\emph{top}) and images (\emph{bottom}). 
  When multiple types have been tried by the authors, we report the one advised.}
  \label{tab:relwork_attention}
\end{table}

\section{Attention Layer Can Implement Convolution Layer}

@@

We take inspiration from \citep[TransformerXL]{dai2019transformerxl} to define relative positional encoding.
They decompose the computation of the attention coefficients as follow:
\begin{align}
    \mathbf{A}_{i, j}^{\mathrm{abs}} &= (\mE_{x_i} + \mU_i)^\top \mW_q^\top \mW_k (\mE_{x_j} + \mU_j)\\
  %
  %
 &=
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{E}_{x_{j}}}_{(a)} 
  +
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{U}_{j}}_{(b)}
  +
  \underbrace{\mathbf{U}_{i}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{E}_{x_{j}}}_{(c)}
  +
  \underbrace{\mathbf{U}_{i}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k} \mathbf{U}_{j}}_{(d)}
\end{align}
with $\mE_{x_i}$ the embedding of token $x_i$, $\mU_i$ the absolute position encoding, $\mR_{i-j}$ relative position encoding. They replace all absolute position encoding with relative ones:
\begin{align}
  \mathbf{A}_{i, j}^{\mathrm{rel}} &=
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k, E} \mathbf{E}_{x_{j}}}_{(a)}
  +
  \underbrace{\mathbf{E}_{x_{i}}^{\top} \mathbf{W}_{q}^{\top} \mathbf{W}_{k, R} {\color{blue} \mathbf{R}_{i-j}}}_{(b)}
  +
  \underbrace{{\color{red} \vu^{\top}} \mathbf{W}_{k, E} \mathbf{E}_{x_{j}}}_{(c)}
  +
  \underbrace{{\color{red} \vv^{\top}} \mathbf{W}_{k, R} {\color{blue} \mathbf{R}_{i-j}}}_{(d)}
  \label{eq:att_rel}
\end{align}

@@

\begin{align}
  f_{\vmu, \mSigma}(\vx) = \exp \left(
    - \frac 1 2
    (\vx - \vmu)^\top \mSigma^{-1} (\vx - \vmu)
  \right)
\end{align}
We drop the normalization factor of the Gaussian distribution as it is replaced by the softmax of the attention mechanism.
The matrix $\mSigma^{-1}$ must be positive semi-definite.
We parametrize it as 
\begin{align}
  \mSigma^{-1} = (\mSigma^{-1/2})^\top\mSigma^{-1/2} = \begin{pmatrix}a & b \\ b & c\end{pmatrix}  
\end{align}
Our goal is to write the exponent as a dot product of two vectors:
\begin{itemize}
  \item one vector $R_\vx$ dependent only on $\vx$, the relative positional encoding,
  \item one vector $\mathbf{t}_h$ dependent only on $\vmu$ and $\mSigma$, the attention head.
\end{itemize}
\begin{align}
  R_\vx &= 
  \begin{pmatrix}
    x_1      & x_2       & x_1      & x_2      & x_1^2   & x_2^2 & x_1 x_2 & 1        & 1         & 1
  \end{pmatrix}\\
  \mathbf{t}_h &= -\frac{1}{2}
  \begin{pmatrix}
    -2a\mu_1 & -2c \mu_2 & -2b\mu_2 & -2b\mu_1 & a       & c     & 2b      & a\mu_1^2 & c\mu_2^2 & 2b\mu_1 \mu_2
  \end{pmatrix}\\
  R_\vx^\top \mathbf{t}_h &=
  -\frac{1}{2} 
  \left[
  a(x_1 - \mu_1)^2 + 2b(x_1-\mu_1)(x_2 -\mu_2) + c(x_2-\mu_2)^2 
  \right]
  = - \frac 1 2 (\vx - \vmu)^\top \mSigma^{-1} (\vx - \vmu)
\end{align}
We can check that setting $b = 0$ and $a=c=\alpha$ recovers the expression of the isotropic Gaussian derived above.

\subsection{Implications to Sinusoid or Learned Positional Encoding}

\subsection{Exploit Attention on Features?} 
\begin{itemize}

@@

\begin{figure}[h]
\includegraphics[width=.16\linewidth]{plots/attention/layer_0.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_1.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_2.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_3.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_4.png}
\includegraphics[width=.16\linewidth]{plots/attention/layer_5.png}
\caption{Contours of attention weights per layer for the 6 layers fully attentional model.}
\end{figure}

\begin{figure}[h]
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_0.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_1.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_2.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_3.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_4.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_5.png}
\includegraphics[width=.13\linewidth]{plots/attention_training/layer_3_6.png}
\caption{Contours of attention weights during training (300 epochs) at layer 3.}
\end{figure}

@@

\section{Discussion}

% \subsubsection*{Author Contributions}
% If you'd like to, you may include  a section for author contributions as is done
% in many journals. This is optional and at the discretion of the authors.

@@

% \appendix
% \section{Appendix}
% You may include other additional sections here.
